\chapter{Shader Local Data: Vertices}

TODO: intro ...

\section{Vertex Data}

In this section we define the vertex data we need to draw our triangle.
We first define the structure of a single vertex.
After that we lay out the vertices of our triangle.

\subsection{Vertex}

Here we define the structure of our vertices.
In this case, our vertex will store its 2D position in normalized device coordinates.
It will also store its color.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={What data we store per vertex},
        label={lst::PositionColorVertex}
        ]{src/ChVertices/PositionColorVertex.cpp}
\end{minipage}

\subsection{Vertex Data}

Since we want to draw a triangle, we specify three vertices.
These vertices are the ones that will be later uploaded to our GPU and be used
for rendering.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={The vertices that our application will use},
        label={lst::PositionColorVertices}
        ]{src/ChVertices/PositionColorVertices.cpp}
\end{minipage}

\section{Shaders}

Since we have changed the structure of our vertices and we also
upload the vertex data to the GPU from our application,
we have to change our vertex shader.
Since our vertices also specify their color, we also change our fragment shader.

\subsection{Vertex Shader}

Our vertex shader takes as input a position value and a color value.
We also want to pass our vertex color to our fragment shader.
For this reason, our vertex shader defines an additional output variable for
the fragment color.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Our new vertex shader},
        label={lst::PositionColorShaderVertex}
        ]{src/ChVertices/PositionColorShader.vert}
\end{minipage}

\subsection{Fragment Shader}

Our fragment shader now takes as input the fragment's color.
We use this value to color our fragment.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Our new fragment shader},
        label={lst::PositionColorShaderFragment}
        ]{src/ChVertices/PositionColorShader.frag}
\end{minipage}

\section{Upload Vertex Data To The GPU}

In order to use our vertex data to render our triangle, we have to upload it
to the GPU for it to be used by our shaders.

\subsection{Understanding The Problem}

Uploading our vertex data to the GPU means that we have to copy it
from the memory that our application uses, to the memory that our GPU uses.
In practice, this means that we have to transfer data from our RAM to our GPU.
Here arises a question: to what kind of GPU memory do we upload
our vertex data?

Modern GPUs have many different types of memories.
Some of these memories fall into two categories.
We have memories that are host visible.
This means that they can be mapped into the application's address space.
Hence they are visible from our application.
We have memories that are device local.
This means that they cannot be mapped to our application's address space.

Host visible memories are slower.
This is due to the fact that they must be visible from both CPU side and GPU side.
This requires particular care from the GPU driver to keep the data consistent.

Device local memories are orders of magnitude faster.
This is due to the fact that they can only be accessed from the GPU.

Keeping in mind that our vertex data doesn't change at run time and that
we use our vertex data every frame, we really would love to use a memory type
that is device local for our vertex buffer.
This would improve performance.
The problem is that we still need to upload our data to the vertex buffer,
and to accomplish this task we have to use a host visible memory type for our
buffer.

\subsection{Our Solution: Idea}

One solution to this problem is to use two buffers.
One buffer, called a staging buffer, will use a host visible memory type.
We use this staging buffer to upload data to the GPU.

We then use another buffer that will use a device local memory type.
After uploading our data to the staging buffer, we issue a memory transfer command
to our GPU.
In this command we say that we want to copy our data from the former buffer to
the latter.
This latter buffer will be the one that our GPU uses as a vertex buffer.

\subsection{How To Create A Buffer}

Before implementing our solution we need to know how to cerate a buffer using Vulkan.
Buffer creation is divided in three steps.
We first create a buffer object.
Then, we allocate our buffer memory.
Finally, we bind our buffer memory to our buffer object.

\subsubsection{Create Buffer Object}

Before creating a buffer object we have to specify its size in bytes, a bitmask
specifying allowed usages of the buffer, and the buffer's sharing mode
when it will be accessed by multiple queue families.
Since all our buffers will only be used by our graphics queue,
we always specify the more performant exclusive sharing mode.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Create a buffer object},
        label={lst::CreateBuffer}
        ]{src/ChVertices/CreateBuffer.cpp}
\end{minipage}

\subsubsection{Allocate Buffer Memory}

Allocating GPU memory requires some additional work.
This is due to the fact that modern GPUs have many different types
of memories.
An added complexity is also caused by the fact that we usually
want our buffer memory to satisfy a given set of memory properties
(host visible, device local, etc ...).

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Allocate buffer memory},
        label={lst::AllocateBufferMemoy}
        ]{src/ChVertices/AllocateBufferMemoy.cpp}
\end{minipage}

We first need to query our buffer memory requirements.
The query's result contain the set of all memory types that are compatible
with our buffer.

With this information at hand, we can pick a memory type which we can
allocate our buffer memory from.
Not only this memory type must be compatible with our buffer.
It also must satisfy the memory properties that we require.
In order to select such memory type, we use an auxiliary function
called \texttt{FindMemoryType}.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Find suitable memory type index},
        label={lst::FindMemoryType}
        ]{src/ChVertices/FindMemoryType.cpp}
\end{minipage}

Now we can finally allocate our memory using \texttt{vkAllocateMemory}.

\subsubsection{Bind Buffer Object And Memory}

We use \texttt{vkBindBufferMemory} to bind the buffer's object and memory together.

\subsection{Our Solution: Implementation}

Now that we know how to allocate a buffer, we can see how our solution
is implemented.

\subsubsection{Create A Staging Buffer}

We use the \texttt{VK\_BUFFER\_USAGE\_TRANSFER\_SRC\_BIT} flag
because our buffer will be used by the GPU as a source for
transfer operations.

We use the \texttt{VK\_MEMORY\_PROPERTY\_HOST\_VISIBLE\_BIT} flag
because we want our buffer to be host visible.

We use the \texttt{VK\_MEMORY\_PROPERTY\_HOST\_COHERENT\_BIT} flag
because we don't want to manually flush our buffer memory.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Crete staging buffer},
        label={lst::CreateStagingBuffer}
        ]{src/ChVertices/CreateStagingBuffer.cpp}
\end{minipage}

\subsubsection{Upload Vertex Data To The Staging Buffer}

Uploading our vertex data to our staging buffer is very simple.
We map the staging buffer memory into our application's address space.
We copy the data.
We unmap the previously mapped memory.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Upload our vertex data to the staging buffer},
        label={lst::MapVertexDataIntoStagingBuffer}
        ]{src/ChVertices/MapVertexDataIntoStagingBuffer.cpp}
\end{minipage}

\subsubsection{Create The Vertex Buffer}

We use the \texttt{VK\_BUFFER\_USAGE\_TRANSFER\_DST\_BIT} flag
because our buffer will be used by the GPU as a destination for
transfer operations.

We use the \texttt{VK\_BUFFER\_USAGE\_VERTEX\_BUFFER\_BIT} flag
because our buffer will be used by the GPU as a source from which
to read vertex data.

We use the \texttt{VK\_MEMORY\_PROPERTY\_DEVICE\_LOCAL\_BIT} flag
because we want our buffer to be device local.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Create vertex buffer},
        label={lst::CreateVertexBuffer}
        ]{src/ChVertices/CreateVertexBuffer.cpp}
\end{minipage}

\subsubsection{Allocate Command Buffer}

Here we simply allocate a command buffer from our graphics command pool.
Is it ok to use the graphics command pool to execute transfer commands?
Yes, because all graphics queues always support transfer commands.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Allocate our transfer command buffer},
        label={lst::AllocateTransferCommandBuffer}
        ]{src/ChVertices/AllocateTransferCommandBuffer.cpp}
\end{minipage}

\subsubsection{Record Copy Command}

Now that we have allocated our command buffer, we can record
our copy command on it.
Here, we are telling our GPU to copy \texttt{copyRegion.size} bytes
from our staging buffer to our vertex buffer.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Record copy command into our command buffer},
        label={lst::RecordCopyCommand}
        ]{src/ChVertices/RecordCopyCommand.cpp}
\end{minipage}

\subsubsection{Submit Command Buffer}

Now we have to submit our command buffer to the GPU graphics queue.
Doing this will start the execution of our copy command.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Submit transfer command buffer},
        label={lst::SubmitCopyCommand}
        ]{src/ChVertices/SubmitCopyCommand.cpp}
\end{minipage}

\subsubsection{Wait For The Command Buffer Execution To Finish}

Here, for simplicity's sake, we wait for our transfer command to finish
before going on with our application.
To do this we call \texttt{vkQueueWaitIdle} on our graphics queue.
This is not a performance issue, since we should be doing this
only during the setup phase of our application.

\subsubsection{Cleanup}

We can now free our transfer command buffer using \texttt{vkFreeCommandBuffers}.
We do this since we won't be using it anymore.
We can also destroy our staging buffer using \texttt{vkFreeMemory}
and \texttt{vkDestroyBuffer}.

\subsection{Review The Process}

Here we can see some pseudocode that outlines all the steps necessary to
create a vertex buffer.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Steps for creating our vertex buffer},
        label={lst::VertexBufferCreation}
        ]{src/ChVertices/VertexBufferSolution.cpp}
\end{minipage}

\section{Pipeline Vertex Input State}

During the creation of our pipeline state object, we must specify
the format of the data that will be passed as input to our pipeline.
To do this we use a \texttt{VkPipelineVertexInputStateCreateInfo} struct.

\texttt{pVertexBindingDescriptions} is an array of vertex input binding descriptions.
\texttt{pVertexAttributeDescriptions} is an array of vertex input attribute descriptions.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Describe the pipeline input data},
        label={lst::VkPipelineVertexInputStateCreateInfo}
        ]{src/ChVertices/VkPipelineVertexInputStateCreateInfo.cpp}
\end{minipage}

\subsection{Binding Descriptions}

A \texttt{VkVertexInputBindingDescription} struct has three fields.
\texttt{binding} is the binding number that this structure describes.
\texttt{stride} is the number of bytes between consecutive elements within the
vertex buffer.
\texttt{inputRate} is a value that specifies whether vertex attribute addressing
is a function of the vertex index or of the instance index.

Some of you may be wondering: what is a vertex input binding?
During rendering, before recording our draw command, we must bind our vertex buffer
in order for our pipeline to use it.
We use \texttt{vkCmdBindVertexBuffers} to do this.
This function takes an array of vertex buffers.
This is because our pipeline can get vertex data from multiple buffers
at the same time.
\texttt{binding} is an index into the \texttt{pBuffers} array bound by
\texttt{vkCmdBindVertexBuffers}.

In our case, all the vertex data is packed together and comes from a single buffer.
Hence, we only have one binding.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Describe our vertex input bindings},
        label={lst::VkVertexInputBindingDescription}
        ]{src/ChVertices/VkVertexInputBindingDescription.cpp}
\end{minipage}

\subsection{Attribute Descriptions}

A vertex attribute is an input variable to a shader that is supplied with
per-vertex (or per-instance) data.
In our case, for each vertex, we supply our shader with a position and a color.
Thus, in our case, our vertex has two attributes: a position attribute and a
color attribute.

Before creating our pipeline state object we must describe all our
vertex attributes.
We do this using an array of \texttt{VkVertexInputAttributeDescription} struct
instances.
\texttt{location} is the shader input location number for this attribute.
It's the one that we use in our vertex shader code.
We have seen it earlier when we wrote \texttt{layout(location = 0)}.
\texttt{binding} is the binding number which this attribute takes its data from.
\texttt{format} is the size and type of the vertex attribute data.
\texttt{offset} is a byte offset of this attribute relative to the start of
an element in the vertex input binding.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Describe our vertex input attributes},
        label={lst::VkVertexInputAttributeDescription}
        ]{src/ChVertices/VkVertexInputAttributeDescription.cpp}
\end{minipage}

\section{Draw Using Our Vertex Data}

The only missing thing now is to tell Vulkan to render a triangle using our
vertex data.
We do this when we record rendering commands into our command buffer.

\begin{minipage}{\linewidth}{\noindent}
    \lstinputlisting[
        language=C++,
        caption={Draw Triangle Using Our Vertex Data},
        label={lst::DrawWithVertexBuffer}
        ]{src/ChVertices/DrawWithVertexBuffer.cpp}
\end{minipage}
